The Power of Two Random Choices

### 2014-01-13 article/article.tex
introduction for applications

\section{Applications}
\label{sec:applications}
\subsection{Hashing}
A simple hash table implementation typically uses a single hash function to map elements to entries in a table. If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a \emph{chain}, usually implemented as a linked list. The maximum time to search for an element is then proportional to the length of the longest chain in the hash table.

@@

\subsection{Online load balancing}
The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers. For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., require the same amount of execution time.

@@

\subsection{Shared memory emulation on distributed memory machines}
The \emph{Parallel Random Access Machine} model\footnote{as defined in the lecture or in~\cite{P03}}, or \emph{PRAM} for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time. It is especially useful for the study of parallel algorithms. However, a PRAM is much harder to realize in hardware than a \emph{Distributed Memory Machine} (DMM). Thus, ways to emulate a PRAM on a DMM have been studied extensivly. Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM. Each memory modul can handle only one memory access at a time step. More accesses to the same module will result in memory contention. The balls-into-bins idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal. For details, see Karp et. al \cite{KLM92} and the papers referenced there.

\subsection{Low congestion circuit routing}
In multi-stage interconnection networks, virtual circuit-switching is used to route communication. For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition). Then, congestions is the maximum number of paths that are routed through a network link. The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.

>>>>

\section{Applications}
\label{sec:applications}
The two-choice paradigm applied to the balls-into-bins model has several useful applications, some of which we will outline here. The applications in \emph{Hashing} (Section \ref{sec:hashing}), \emph{Online Load Balancing} (Section \ref{sec:loadbalancing}) and \emph{Emulation of shared memory machines on distributed memory machines} \ref{sec:DMM}) share a close relation to load balancing. Section \ref{sec:circuitrouting} applies the balls-into-bins model and the two-choice paradigm in a different context, namely that of routing circuits in multi-stage interconnection networks with minimal congestion.

\subsection{Hashing}
\label{sec:hashing}
A simple hash table implementation typically uses a single hash function to map elements to entries in a table. If more than one element is mapped to the same entry, all elements mapped to this table entry are stored in a \emph{chain}, usually implemented as a linked list. The maximum time to search for an element is then proportional to the length of the longest chain in the hash table.

@@

\subsection{Online load balancing}
\label{sec:loadbalancing}
The following simple scenario is a typical application of online load balancing: Consider $n$ servers\footnote{Such as file-, database- or network servers} and $m$ requests issued by clients that need to be handled by these servers. For simplicity, assume that each task can be handled by any server and that all tasks are of the same size, i.e., require the same amount of execution time.

@@

\subsection{Emulation of shared memory machines on distributed memory machines}
\label{sec:DMM}
The \emph{Parallel Random Access Machine} model\footnote{as defined in the lecture or in~\cite{P03}}, or \emph{PRAM} for short, is a high level abstraction of a shared memory machine, in which memory access by any processor to any memory location is assumed to require constant time. It is especially useful for the study of parallel algorithms. However, a PRAM is much harder to realize in hardware than a \emph{Distributed Memory Machine} (DMM). Thus, ways to emulate a PRAM on a DMM have been studied extensivly. Such an emulation distributes the processors and memory cells of the PRAM to the processors and memory modules (typically one per processor) of the DMM with the goal to reduce the slowdown of the emulation, which is the time the DMM requires to simulate one step of the PRAM. Each memory modul can handle only one memory access at a time step. More accesses to the same module will result in memory contention. The balls-into-bins idea is useful for managing communication between the processors and the memory modules such that memory contention is minimal. For details, see Karp et. al \cite{KLM92} and the papers referenced there.

\subsection{Low congestion circuit routing}
\label{sec:circuitrouting}
In multi-stage interconnection networks, virtual circuit-switching is used to route communication. For simplicity, we consider the \emph{permutation routing problem}, where one request originates at each of the $n$ input nodes and one request is destined for each of the $n$ output nodes and assume a \emph{butterfly network} as the underlying interconnection network (see \cite{CLR09} for a detailed definition). Then, congestions is the maximum number of paths that are routed through a network link. The goal of a \emph{circuit routing algorithm} is to to allocate paths for all $n$ communication requests such that congestion is minimal.
